```{r envSetup, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE,
    out.width = "65%",
    fig.align = "center",
    comment = ">"
    )
```



```{r, include = FALSE, echo = FALSE}
#library(tidyverse)
library(corrr)
library(psych)
library(lavaan)
#library(kableExtra)
library(dplyr)
library(tidyr)
library(ggplot2)
library(survey)
library(haven)
library(rempsyc)
library(broom)
library(report)
library(effectsize)
library(aod)
library(readr)
#library(tidymodels)
library(stargazer)
library(forcats)
library(ggcorrplot)
library(caret)
library(knitr)
library(ROCR)
library(jtools)
#=library(car)
library(xtable)
#library(texreg)
#library(svyVGAM)
library(glmnet)
library(ggpubr)
library(lme4)
library(nlme)
library(weights)
library(miscTools)
library(systemfit)
library(multcomp)
require(ggplot2)
require(GGally)
require(reshape2)
#require(compiler)
#require(parallel)
#require(boot)
require(lattice)
library(HLMdiag)
#library(DHARMa)
library(car) #for the Levene test which we will not discuss here
library(Matrix)
library(brms)
library(margins)
library(performance)
library(ggnewscale)
library(ggeffects)
library(bayestestR) # for hypothesis testing
library(brmsmargins)
library(ggeffects)
library(marginaleffects)
library(effects)
library(margins)
library(modelr)
library(plm)
library(effectsize)
library(aod)
library(readr)
library(tidymodels)
library(stargazer)
library(forcats)
library(ggcorrplot)
library(caret)
library(knitr)
library(ROCR)
library(jtools)
#=library(car)
library(xtable)
#library(texreg)
#library(svyVGAM)
library(glmnet)
library(ggpubr)
library(foreign)
library(AER)
library(lme4)
library(brms)
library(glmnet)
library(ggpubr)
library(formatR)
library(pglm)
library(acqr)
library(readr)
library(tidymodels)
library(foreign)
library(NLP)
library(tm) # text mining 
library(tidytext) # text analysis 
#library(janeaustenr)
library(stringr)
library(igraph) # manipulating and analyzing networks
library(ggraph)
library(igraph)
library(widyr)
#library(tm.plugin.webmining)
library(purrr)
library(topicmodels)
library(scales)
#library(mallet) 
library(cld2)
library(readxl)
library(reshape2)
library(pheatmap)
library(stopwords)
library(gridExtra)
library(grid)
library(reshape2)
library(igraph)
library(treemapify)
library(networkD3)
library(tidyr)
library(ggeffects)
library(sandwich)
library(lmtest)
library(pscl)
library(texreg)
library(geepack)
```


```{r}

#df_00_wa <- readRDS("wa_statewide.rds")
#df_00_alb <- readRDS("albany.rds")
#df_00_ca <- readRDS("ca_statewide.rds")
#df_00_chi <- readRDS("chicago.rds")
df_00_fl <- readRDS("fl_statewide.rds")
#df_00_NY <- readRDS("ny_statewide.rds")
#df_00_LA <- readRDS("los_angeles.rds")
#df_00_SND <- readRDS("san_diego.rds")
#df_00_SF <- readRDS("san_francisco.rds")
#df_00_SEAT <- readRDS("seattle.rds")
#df_00_ILL <- readRDS("il_statewide.rds")


```


```{r}
colnames(df_00_wa)
```

```{r}
colnames(df_00_alb)
```

```{r}

colnames(df_00_ca)
```

```{r}
colnames(df_00_chi)

```


```{r}
colnames(df_00_fl)
```


```{r}
#colnames(df_00_NY)
```


```{r}
#colnames(df_00_LA)
```


```{r}
#colnames(df_00_SND)
```


```{r}
#colnames(df_00_SF)
```


```{r}
#colnames(df_00_SEAT)
```

```{r}
#colnames(df_00_ILL)

```
The only ones that have the search_conducted are: 
- df_00_ILL
- df_00_SF
- df_00_ca
- df_00_SND
- df_00_fl
- df_00_wa

```{r}
main_df <- df_00_fl
df <- main_df
```


```{r}
glimpse(df)
```


```{r}
df %>% filter(!is.na(reason_for_search)) %>% dplyr::select(reason_for_search)
```

```{r}
df2 <- df %>% dplyr::select(date, time, location, county_name, department_name, subject_age, subject_race, 
                            subject_sex, officer_age, officer_race, officer_sex, officer_years_of_service, type, violation, arrest_made, outcome, search_conducted)
```

```{r}
dim(df2)
```

```{r}
df2_no_na <- df2 %>% filter(if_all(everything(), ~ !is.na(.)))

```

```{r}
dim(df2_no_na)
```
```{r}
df2_clean <- df2_no_na
```


```{r}
describe_df <- psych::describe(df2_clean %>% dplyr::select(where(is.numeric)))

```

```{r}
describe_df_unclean <- psych::describe(df2 %>% dplyr::select(where(is.numeric)))

```

```{r}
as.data.frame(describe_df) %>% dplyr::select(n, mean, sd, median, min, max, skew)
```





```{r}
xtable(as.data.frame(describe_df_unclean))
```

```{r}
categorical_summary <- function(data) {
  data %>%
    dplyr::select(where(~ is.factor(.) || is.character(.))) %>%
    purrr::map_dfr(
      ~ as.data.frame(table(.x, useNA = "ifany")) %>%
        dplyr::mutate(Percent = round(100 * Freq / sum(Freq), 2)),
      .id = "Variable"
    )
}

```


```{r}
paste("Number of unique dates in dataset are ", n_distinct(df2$date), ".")
paste("Number of unique times in dataset are ", n_distinct(df2$time), ".")
paste("Number of unique locations in dataset are ", n_distinct(df2$location), ".")
paste("Number of unique counties in dataset are ", n_distinct(df2$county_name), ".")
paste("Number of unique departments in dataset are ", n_distinct(df2$department_name), ".")
paste("Number of unique subject ages in dataset are ", n_distinct(df2$subject_age), ".")
paste("Number of unique subject races in dataset are ", n_distinct(df2$subject_race), ".")
paste("Number of unique subject genders in dataset are ", n_distinct(df2$subject_sex), ".")
paste("Number of unique officer ages in dataset are ", n_distinct(df2$officer_age), ".")
paste("Number of unique officer races in dataset are ", n_distinct(df2$officer_race), ".")
paste("Number of unique officer genders in dataset are ", n_distinct(df2$officer_sex), ".")
paste("Number of unique officer years of service in dataset are ", n_distinct(df2$officer_years_of_service), ".")
paste("Number of unique types of stop in dataset are ", n_distinct(df2$type), ".")
paste("Number of unique violation in dataset are ", n_distinct(df2$violation), ".")
paste("Number of unique arrest results in dataset are ", n_distinct(df2$arrest_made), ".")
paste("Number of unique outcomes in dataset are ", n_distinct(df2$outcome), ".")
paste("Number of unique search binary in dataset are ", n_distinct(df2$search_conducted), ".")


```
```{r}
paste("Number of unique dates in dataset are ", n_distinct(df2_clean$date), ".")
paste("Number of unique times in dataset are ", n_distinct(df2_clean$time), ".")
paste("Number of unique locations in dataset are ", n_distinct(df2_clean$location), ".")
paste("Number of unique counties in dataset are ", n_distinct(df2_clean$county_name), ".")
paste("Number of unique departments in dataset are ", n_distinct(df2_clean$department_name), ".")
paste("Number of unique subject ages in dataset are ", n_distinct(df2_clean$subject_age), ".")
paste("Number of unique subject races in dataset are ", n_distinct(df2_clean$subject_race), ".")
paste("Number of unique subject genders in dataset are ", n_distinct(df2_clean$subject_sex), ".")
paste("Number of unique officer ages in dataset are ", n_distinct(df2_clean$officer_age), ".")
paste("Number of unique officer races in dataset are ", n_distinct(df2_clean$officer_race), ".")
paste("Number of unique officer genders in dataset are ", n_distinct(df2_clean$officer_sex), ".")
paste("Number of unique officer years of service in dataset are ", n_distinct(df2_clean$officer_years_of_service), ".")
paste("Number of unique types of stop in dataset are ", n_distinct(df2_clean$type), ".")
paste("Number of unique violation in dataset are ", n_distinct(df2_clean$violation), ".")
paste("Number of unique arrest results in dataset are ", n_distinct(df2_clean$arrest_made), ".")
paste("Number of unique outcomes in dataset are ", n_distinct(df2_clean$outcome), ".")
paste("Number of unique search binary in dataset are ", n_distinct(df2_clean$search_conducted), ".")

```

```{r}
df3 <- df2_clean %>% dplyr::select(subject_race, department_name, subject_sex, officer_race, officer_sex, arrest_made, outcome, search_conducted)
```

```{r}
xtable(categorical_summary(df3))
```

```{r}
time_summary <- function(data) {
  data %>%
    dplyr::select(where(~ inherits(., "Date") || inherits(., "POSIXct"))) %>%
    purrr::imap_dfr(~ {
      tibble::tibble(
        Variable = .y,
        Min = min(.x, na.rm = TRUE),
        Max = max(.x, na.rm = TRUE),
        Range = difftime(max(.x, na.rm = TRUE), min(.x, na.rm = TRUE), units = "days"),
        Missing = sum(is.na(.x)),
        Total = length(.x)
      )
    })
}

```

```{r}
summary_date <- tibble::tibble(
  Min = min(df2_clean$date, na.rm = TRUE),
  Max = max(df2_clean$date, na.rm = TRUE),
  Range = difftime(max(df2_clean$date, na.rm = TRUE), min(df2_clean$date, na.rm = TRUE), units = "days"),
  Missing = sum(is.na(df2_clean$date)),
  Total = length(df2_clean$date)
)
summary_date
```

```{r}
df$datetime <- as.POSIXct(paste(df$date, df$time), format = "%Y-%m-%d %H:%M:%S")
```

```{r}
df2_clean$datetime <- as.POSIXct(paste(df2_clean$date, df2_clean$time), format = "%Y-%m-%d %H:%M:%S")

```


```{r}
summary_date2 <- tibble::tibble(
  Min = min(df$datetime, na.rm = TRUE),
  Max = max(df$datetime, na.rm = TRUE),
  Range = difftime(max(df$datetime, na.rm = TRUE), min(df$datetime, na.rm = TRUE), units = "days"),
  Missing = sum(is.na(df$datetime)),
  Total = length(df$datetime)
)
summary_date2
```

```{r}
summary_date2 <- tibble::tibble(
  Min = min(df2_clean$datetime, na.rm = TRUE),
  Max = max(df2_clean$datetime, na.rm = TRUE),
  Range = difftime(max(df2_clean$datetime, na.rm = TRUE), min(df2_clean$datetime, na.rm = TRUE), units = "days"),
  Missing = sum(is.na(df2_clean$datetime)),
  Total = length(df2_clean$datetime)
)
summary_date2
```



df <- df %>% mutate(
    date = as.numeric(date),
    id = as.factor(as.integer(factor(objectid)))
)


non_unique_id <- df %>% count(id) %>% 
    filter(n > 1) %>% pull(id)

head(df %>% filter(id %in% non_unique_id))

sum(is.na(df$date))



```{r fig.height=15, fig.width=20}
corM <- Hmisc::rcorr(as.matrix(
    df2_clean %>% dplyr::select_if(~ is.numeric(.))
))

```

```{r}
reg_corM <- as.data.frame(corM$r)
reg_corM <- reg_corM %>% dplyr::select(where(~ !all(is.na(.))))

```

```{r}
#reg_corM %>% dplyr::select(MBANK) %>% arrange(desc(abs(MBANK)))
```


```{r}
#rownames(heatmap_data) <- FACandNSFAC_pivot$FAC
corM <- Hmisc::rcorr(as.matrix(
    df2_clean %>% dplyr::select_if(is.numeric)
))

# Extract the correlation values (correlation matrix)
reg_corM <- as.data.frame(corM$r)

# Step 2: Remove columns and rows with all NA values
reg_corM <- reg_corM %>%
  dplyr::select(where(~ !all(is.na(.)))) %>%
  filter(rowSums(is.na(.)) < ncol(.))

# Step 3: Convert to a numeric matrix and replace NA with 0 (optional)
data <- as.matrix(reg_corM)
data[is.na(data)] <- 0  # Replace NA with 0 if needed

# Step 4: Create a heatmap using pheatmap
pheatmap(data,
         main = "Heatmap of Correlation Matrix",
         display_numbers = TRUE,  # Show correlation values
         clustering_distance_rows = "euclidean",  # Distance metric for clustering
         clustering_distance_cols = "euclidean",
         clustering_method = "complete",  # Hierarchical clustering method
         color = colorRampPalette(c("cyan", "white", "orange"))(50))  # Customize color palette


```

```{r}
cor_values <- as.data.frame(corM$r)   
p_values <- as.data.frame(corM$P) 


```

```{r}
p_values
```

```{r}
searched <- df2_clean %>% filter(search_conducted == "TRUE")
```

```{r}
ggplot(searched, aes(x = subject_race)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Searches by Race",
       x = "Race",
       y = "Number of Searches") +
  theme_minimal()
```

```{r, fig.height=4, fig.width=10}

sbjct_race_search <- df2_clean %>%
      count(subject_race, search_conducted) %>%
      ggplot(aes(x = as.factor(search_conducted), y = n, fill = subject_race)) +
      geom_bar(stat = "identity", position = "fill") +
      labs(fill = "Race", x = "Search Conducted", y = "Proportion", title = "Subject Race & Searches") +
      scale_fill_brewer(palette = "Set2") +
      theme_minimal()

officer_race_search <- df2_clean %>%
      count(officer_race, search_conducted) %>%
      ggplot(aes(x = as.factor(search_conducted), y = n, fill = officer_race)) +
      geom_bar(stat = "identity", position = "fill") +
      labs(fill = "Race", x = "Search Conducted", y = "Proportion", title = "Officer Race & Searches") +
      scale_fill_brewer(palette = "Set1") +
      theme_minimal()

ggarrange(sbjct_race_search, officer_race_search, ncol = 2)
```



```{r, fig.height=4, fig.width=10}
sex_subject_search <- df2_clean %>%
      count(subject_sex, search_conducted) %>%
      ggplot(aes(x = as.factor(search_conducted), y = n, fill = subject_sex)) +
      geom_bar(stat = "identity", position = "fill") +
      labs(fill = "Gender", x = "Search Conducted", y = "Proportion", title = "Subject Sex & Searches") +
      scale_fill_brewer(palette = "Set2") +
      theme_minimal()


sex_officer_search <- df2_clean %>%
      count(officer_sex, search_conducted) %>%
      ggplot(aes(x = as.factor(search_conducted), y = n, fill = officer_sex)) +
      geom_bar(stat = "identity", position = "fill") +
      labs(fill = "Gender", x = "Search Conducted", y = "Proportion", title = "Officer Sex & Searches") +
      scale_fill_brewer(palette = "Set3") +
      theme_minimal()

ggarrange(sex_subject_search, sex_officer_search, ncol = 2)

```

```{r}

```

```{r}
df2_clean %>%
  filter(search_conducted == 1) %>%
  mutate(age_group = cut(subject_age, breaks = seq(0, 100, by = 10), right = FALSE)) %>%
  count(age_group) %>%
  ggplot(aes(x = age_group, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Number of Searches by Age Group",
       x = "Age Group",
       y = "Number of Searches") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, fig.height=4, fig.width=10}
sbjct_age_search <- df2_clean %>%
      filter(search_conducted == 1) %>%
      count(subject_age) %>%
      ggplot(aes(x = subject_age, y = n)) +
      geom_line(color = "steelblue", size = 1) +
      labs(title = "Number of Searches by Subject Age",
           x = "Age (Subject)",
           y = "Number of Searches") +
      theme_minimal()

officer_age_search <- df2_clean %>%
      filter(search_conducted == 1) %>%
      count(officer_age) %>%
      ggplot(aes(x = officer_age, y = n)) +
      geom_line(color = "darkred", size = 1) +
      labs(title = "Number of Searches by Officer Age",
           x = "Age (Officer)",
           y = "Number of Searches") +
      theme_minimal()

ggarrange(sbjct_age_search, officer_age_search, ncol = 2)
```

```{r}
df2_clean %>%
      filter(search_conducted == 1) %>%
      count(officer_years_of_service) %>%
      ggplot(aes(x = officer_years_of_service, y = n)) +
      geom_line(color = "darkgreen", size = 1) +
      labs(title = "Number of Searches based on Officer's years of service",
           x = "Officer Years of Service",
           y = "Number of Searches") +
      theme_minimal()
```
```{r}
n_distinct(df2_clean$officer_years_of_service)
```



```{r, fig.width=10, fig.height=5}
# gg_smartwatch_onbanking_weighted <- ggplot(ds_smwtch_cln, aes(x = CNSMR_TYPE, y = scaled_wtpg, fill = as.factor(MBANK))) +
#     geom_bar(stat = "identity", position = "stack") +
#   labs(fill = "M-Bank", x = "User Type", y = "Weighted Proportion") + 
#   fill_palette("Pastel") 
# 
# gg_smartwatch_onbanking_Notweighted <- ggplot(ds_smwtch_cln, aes(x = CNSMR_TYPE, fill = as.factor(MBANK))) +
#   geom_bar(position = "fill") +
#   labs(fill = "M-Bank", x = "User Type", y = "Proportion") + 
#   fill_palette("Pastel") 
# 
#     
# ggarrange(gg_smartwatch_onbanking_weighted, gg_smartwatch_onbanking_Notweighted, ncol = 2)
```



corM <- Hmisc::rcorr(as.matrix(
    ds_smphn_cln %>% dplyr::select_if(is.numeric)
))

# Extract the correlation values (correlation matrix)
 

result_good <- cor_values %>% dplyr::select(where(~ !all(is.na(.)))) %>%
  filter(rowSums(is.na(.)) < ncol(.)) %>% 
  dplyr::select(MBANK) %>%  
  mutate(variable = rownames(cor_values),  # Add variable names for easier reference
         p_value = p_values$MBANK) %>%   # Add p-values for ONBANK
  filter(abs(MBANK) > 0.1) %>%           # Filter correlations > threshold
  arrange(desc(abs(MBANK)))              # Sort by absolute correlation values

result_bad <- cor_values %>% dplyr::select(where(~ !all(is.na(.)))) %>%
  filter(rowSums(is.na(.)) < ncol(.)) %>% 
  dplyr::select(MBANK) %>%  
  mutate(variable = rownames(cor_values),  # Add variable names for easier reference
         p_value = p_values$MBANK) %>%   # Add p-values for ONBANK
  filter(abs(MBANK) < 0.05) %>%           # Filter correlations > threshold
  arrange(desc(abs(MBANK)))





### TEST - Chi-Squared

```{r}
ctab <- table(df2_clean$subject_race, df2_clean$search_conducted)
ctab
```
```{r}
ctab <- ctab[rowSums(ctab) > 0, colSums(ctab) > 0]

chisq.test(ctab)
```


ctab <- table(df3$, df3$)
ctab














